---
title: 'Project Assignment #1 - Reproducible Research'
output: html_document
This is the project assignment for Data science course: reproducible resuearch. 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

First, the following codes will unzip the activity data and read them in a dataframe:Act. Transform the date into date formate

```{r, results='hide'}
library(lubridate)
unzip("activity.zip")
Act <- read.csv("./activity.csv")
Act$date <- ymd(Act$date)
```

Sum the total steps each day by groub_by function and summarise function. The new datatable was saved in Act_stepsum. A histogram was plotted for Act_stepsum

```{r}
library(dplyr)
Act_stepsum <- summarise(group_by(Act, date), steps = sum(steps, na.rm = T))
with(Act_stepsum, hist(steps, breaks = seq(from=0, to=25000, by=2500), xlab = "total steps", ylab = "number of days", main = "Histogram of total number of steps taken each day", col = "blue"))
```

Next step, I will calculate the mean and median number of steps taken each day and store them in Act_steps_mean and Act_steps_median

```{r}
Act_steps_mean <- round(mean(Act_stepsum$steps, na.rm = TRUE))
Act_steps_median <- median(Act_stepsum$steps)
print(paste("the mean of daily steps is", Act_steps_mean))
print(paste("the median of daily steps is", Act_steps_median))
```

Then we will plot the average of steps taken in the interval
```{r}
library(ggplot2)
Act_steps_avg <- summarise(group_by(Act, interval), steps = mean(steps, na.rm = TRUE))
ggplot(data = Act_steps_avg, aes(x = interval, y = steps))+
        geom_line()+
        xlab("5 minutes interval")+
        ylab("averaged steps")+
        ggtitle("Time series: averaged steps in 5 minutes interval")
```

Then we will find out the maximum number of steps and the interval
```{r}
max_index <- which.max(Act_steps_avg$steps)
max_interval <- Act_steps_avg[max_index, 1]
max_steps <- round(Act_steps_avg[max_index, 2])
print(paste("the max interval happens at", max_interval))
print(paste("the max steps is", max_steps))
```

We will umpute the missing value in the dataset
```{r}
Total_NA <- sum(is.na(Act$steps))
print(paste("the total missing rows are", Total_NA))
Act <- mutate(Act, steps_fillNA = steps)

#the following codes used average steps per interval to replace the NA, stored in a new column called "steps_fillNA"

for (i in 1:nrow(Act)){
        if (is.na(Act[i,4])){
                Act[i,4] <- Act_steps_avg[Act_steps_avg$interval == Act[i,3],2]
        }
}
print(paste("the total NA after imputing is", sum(is.na(Act$steps_fillNA))))

#generating the new sums/mean/median

library(gridExtra)
Act_stepsum_new <- summarise(group_by(Act, date), steps = sum(steps_fillNA))
plot1 <- qplot(Act_stepsum$steps, binwidth = 1000, ylim = c(0,15), main = "original", xlab = "total steps taken each day")
plot2 <- qplot(Act_stepsum_new$steps, binwidth = 1000, ylim = c(0,15), main = "After imputing NAs", xlab = "total steps taken each day")
grid.arrange(plot1, plot2, ncol = 2)

#computing the new mean and median
mean_new <- round(mean(Act_stepsum_new$steps))
median_new <- round(median(Act_stepsum_new$steps))
print(paste("the new mean is", mean_new))
print(paste("the new median is", median_new))
```

Next let's compare the average steps taken per 5-minutes interval between weekdays and weekends
```{r}
#first let's separate the workdays and weekends
weekends <- c("Saturday", "Sunday")
Act_workdays <- filter(Act, weekdays(date) != weekends)
Act_weekends <- filter(Act, weekdays(date) == weekends)
#Then we will calculate the average steps at each interval
Act_workdays_avg <- summarise(group_by(Act_workdays, interval), steps = mean(steps_fillNA))
Act_weekends_avg <- summarise(group_by(Act_weekends, interval), steps = mean(steps_fillNA))
#next we will plot the time series for each of them
plot1 <- ggplot(data = Act_workdays_avg, aes(x = interval, y = steps))+
        geom_line()+
        ylim(0,250)+
        ggtitle("Averaged steps in workdays")+
        xlab("5 minutes interval")+
        ylab("averaged steps")
plot2 <- ggplot(data = Act_weekends_avg, aes(x = interval, y = steps))+
        geom_line()+
        ylim(0,250)+
        ggtitle("Averaged steps in weekends")+
        xlab("5 minutes interval")+
        ylab("averaged steps")
grid.arrange(plot1, plot2, ncol = 2)
```

